---
title: "writeup"
output: github_document
---

# Coverage probability

Coverage probability is an important operating characteristic of methods for constructing interval estimates, particularly confidence intervals.

Definition: For the purposes of this deliverable, define the 95% confidence interval of the median to be the middle 95% of sampling distribution of the median. Similarly, the 95% confidence interval of the mean, standard deviation, etc. is the middle 95% of the respective sampling distribution.

```{r}
library(tidyverse)
library(stats4)
```

## step 1

Generate a single sample from a standard normal distribution of size N = 201. Explain to the reader how you use MLE to estimate the distribution.

```{r}
set.seed(111)
single_sample = rnorm(201, 0, 1)
fun <- function(a, b)
  sum(-log(dnorm(single_sample, a, b)))
z <- mle(minuslogl = fun, start = list(a = 0, b = 1))
mean = coef(z)[1]
sd = coef(z)[2]
mean
sd
```

## step 2

Show the reader how you approximate the sampling distribution of the median, conditional on the estimate of the distribution in the previous step.

```{r}
median_sample <- c()
for (i in 1:100) {
  x = median(rnorm(1000, mean, sd))
  median_sample <- c(median_sample, x)
}
hist(median_sample)
```

## step 3

Describe how you calculate a 95% confidence interval from the approximated sampling distribution.

```{r}
ci = quantile(median_sample, probs = c(0.025, 0.975))
ci
```

## step 4

Explain the concept of coverage probability. Explain your code for calculating the coverage probability.

```{r}
coverage <- c()
for (i in 1:1000) {
  data <- rnorm(201, 0, 1)
  
  mean = mean(data)
  sd = sqrt(length(data-1)/length(data)*var(data))
  
  median_sample <- c()
  for (i in 1:100) {
  x = median(rnorm(201, mean, sd))
  median_sample <- c(median_sample, x)
  }
  
  ci = quantile(median_sample, probs = c(0.025, 0.975))
  
  coverage <- c(coverage, ci[1] <= 0 & ci[2] >= 0)
}

mean(coverage)
```

```{r}
coverage <- c()
for (i in 1:1000) {
  data <- rnorm(201, 0, 1)
  
  median_sample <- c()
  for (i in 1:100) {
  x = median(sample(data, size = 201, replace = TRUE))
  median_sample <- c(median_sample, x)
  }
  
  ci = quantile(median_sample, probs = c(0.025, 0.975))
  
  coverage <- c(coverage, ci[1] <= 0 & ci[2] >= 0)
}

mean(coverage)

```
