---
title: "writeup"
output:
  html_document:
    df_print: paged
---

# 07-mle-and-mm

Maximum likelihood (MLE) and method of moments (MM) are two common methods for constructing a model. For this assignment, we are going to model the unknown distributions with maximum likelihood and method of moments.

## Data and preparation

```{r message=FALSE}
library(tidyverse)
require(dplyr)
library(stats4)
Hmisc::getHdata(nhgh)
d1 <- nhgh %>% 
  filter(sex == "female") %>% 
  filter(age >= 18) %>% 
  select(gh, ht) %>% 
  filter(1:n()<=1000)
```

## Glycohemoglobin (MLE)

### Normal distribution

In this part, we assume that the Glycohemoglobin data is sampled from the normal distribution, and we plan to calculate the parameters first. 

```{r}
gh <- d1$gh
fun <- function(a, b)
  sum(-log(dnorm(gh, a, b)))
z <- mle(minuslogl = fun, start = list(a = 6, b = 1))
a <- coef(z)[1]
b <- coef(z)[2]
a
b
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(gh, breaks = 20, freq = FALSE)
curve(dnorm(x, a, b), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(gh), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pnorm(x, a, b), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- gh
q_candidate <- qnorm

x <- q_candidate((1:200)/200, a, b)
y <- quantile(gh, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qnorm(0.5, a, b)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rnorm(1e4, a, b)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Gamma distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Gamma distribution, and we plan to calculate the parameters first. 

```{r}
gh <- d1$gh
fun <- function(a, b)
  sum(-log(dgamma(gh, a, b)))
z <- mle(minuslogl = fun, start = list(a = 40, b = 7))
a <- coef(z)[1]
b <- coef(z)[2]
a
b
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(gh, breaks = 20, freq = FALSE)
curve(dgamma(x, a, b), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(gh), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pgamma(x, a, b), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- gh
q_candidate <- qgamma

x <- q_candidate((1:200)/200, a, b)
y <- quantile(gh, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qgamma(0.5, a, b)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rgamma(1e4, a, b)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Weibull distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Weibull distribution, and we plan to calculate the parameters first. 

```{r}
gh <- d1$gh
fun <- function(a, b)
  sum(-log(dweibull(gh, a, b)))
z <- mle(minuslogl = fun, start = list(a = 4, b = 6))
a <- coef(z)[1]
b <- coef(z)[2]
a
b
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(gh, breaks = 20, freq = FALSE)
curve(dweibull(x, a, b), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(gh), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pweibull(x, a, b), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- gh
q_candidate <- qweibull

x <- q_candidate((1:200)/200, a, b)
y <- quantile(gh, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x,y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qweibull(0.5, a, b)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rweibull(1e4, a, b)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

## Glycohemoglobin (MM)

### Normal distribution

In this part, we assume that the Glycohemoglobin data is sampled from the normal distribution, and we plan to calculate the parameters first. 

```{r}
gh <- d1$gh
xb <- mean(gh)
s2 <- var(gh)
mu <- xb
sigma <- sqrt(s2)
mu
sigma
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(gh, breaks = 20, freq = FALSE)
curve(dnorm(x, mu, sigma), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(gh), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pnorm(x, mu, sigma), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- gh
q_candidate <- qnorm

x <- q_candidate((1:200)/200, mu, sigma)
y <- quantile(gh, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qnorm(0.5, mu, sigma)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rnorm(1e4, mu, sigma)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Gamma distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Gamma distribution, and we plan to calculate the parameters first. 

```{r}
gh <- d1$gh
xb <- mean(gh)
s2 <- var(gh)
lh <- xb/s2
ch <- xb^2/s2
lh
ch
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(gh, breaks = 20, freq = FALSE)
curve(dgamma(x, ch, lh), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(gh), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pgamma(x, ch, lh), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- gh
q_candidate <- qgamma

x <- q_candidate((1:200)/200, ch, lh)
y <- quantile(gh, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qgamma(0.5, ch, lh)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rgamma(1e4, ch, lh)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Weibull distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Weibull distribution, and we plan to calculate the parameters first. 

```{r}
gh <- d1$gh
xb <- mean(gh)
s2 <- var(gh)
cv2 = s2/xb^2
fun <- function(beta, cv2){
  abs(gamma(1 + 2/beta)/(gamma(1 + 1/beta))^2 - (1 + cv2))
}
k = optimize(fun, interval = c(0, 100), cv2 = cv2, maximum = FALSE)$minimum
lambda = xb/gamma(1 + 1/k)
k
lambda
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(gh, breaks = 20, freq = FALSE)
curve(dweibull(x, k, lambda), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(gh), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pweibull(x, k, lambda), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- gh
q_candidate <- qweibull

x <- q_candidate((1:200)/200, k, lambda)
y <- quantile(gh, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x,y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qweibull(0.5, k, lambda)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rweibull(1e4, k, lambda)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

## Height of adult females (MLE)

### Normal distribution

In this part, we assume that the Glycohemoglobin data is sampled from the normal distribution, and we plan to calculate the parameters first. 

```{r}
ht <- d1$ht
fun <- function(a, b)
  sum(-log(dnorm(ht, a, b)))
z <- mle(minuslogl = fun, start = list(a = 160, b = 10))
a <- coef(z)[1]
b <- coef(z)[2]
a
b
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(ht, breaks = 20, freq = FALSE)
curve(dnorm(x, a, b), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(ht), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pnorm(x, a, b), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- ht
q_candidate <- qnorm

x <- q_candidate((1:200)/200, a, b)
y <- quantile(ht, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qnorm(0.5, a, b)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rnorm(1e4, a, b)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Gamma distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Gamma distribution, and we plan to calculate the parameters first. 

```{r}
ht <- d1$ht
fun <- function(a, b)
  sum(-log(dgamma(ht, a, b)))
z <- mle(minuslogl = fun, start = list(a = 450, b = 2))
a <- coef(z)[1]
b <- coef(z)[2]
a
b
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(ht, breaks = 20, freq = FALSE)
curve(dgamma(x, a, b), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(ht), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pgamma(x, a, b), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- ht
q_candidate <- qgamma

x <- q_candidate((1:200)/200, a, b)
y <- quantile(ht, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qgamma(0.5, a, b)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rgamma(1e4, a, b)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Weibull distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Weibull distribution, and we plan to calculate the parameters first. 

```{r}
ht <- d1$ht
fun <- function(a, b)
  sum(-log(dweibull(ht, a, b)))
z <- mle(minuslogl = fun, start = list(a = 20, b = 160))
a <- coef(z)[1]
b <- coef(z)[2]
a
b
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(ht, breaks = 20, freq = FALSE)
curve(dweibull(x, a, b), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(ht), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pweibull(x, a, b), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- ht
q_candidate <- qweibull

x <- q_candidate((1:200)/200, a, b)
y <- quantile(ht, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x,y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qweibull(0.5, a, b)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rweibull(1e4, a, b)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

## Glycohemoglobin (MM)

### Normal distribution

In this part, we assume that the Glycohemoglobin data is sampled from the normal distribution, and we plan to calculate the parameters first. 

```{r}
ht <- d1$ht
xb <- mean(ht)
s2 <- var(ht)
mu <- xb
sigma <- sqrt(s2)
mu
sigma
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(ht, breaks = 20, freq = FALSE)
curve(dnorm(x, mu, sigma), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(ht), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pnorm(x, mu, sigma), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- ht
q_candidate <- qnorm

x <- q_candidate((1:200)/200, mu, sigma)
y <- quantile(ht, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qnorm(0.5, mu, sigma)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rnorm(1e4, mu, sigma)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Gamma distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Gamma distribution, and we plan to calculate the parameters first. 

```{r}
ht <- d1$ht
xb <- mean(ht)
s2 <- var(ht)
lh <- xb/s2
ch <- xb^2/s2
lh
ch
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(ht, breaks = 20, freq = FALSE)
curve(dgamma(x, ch, lh), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(ht), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pgamma(x, ch, lh), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- ht
q_candidate <- qgamma

x <- q_candidate((1:200)/200, ch, lh)
y <- quantile(ht, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x, y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qgamma(0.5, ch, lh)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rgamma(1e4, ch, lh)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

### Weibull distribution

In this part, we assume that the Glycohemoglobin data is sampled from the Weibull distribution, and we plan to calculate the parameters first. 

```{r}
ht <- d1$ht
xb <- mean(ht)
s2 <- var(ht)
cv2 = s2/xb^2
fun <- function(beta, cv2){
  abs(gamma(1 + 2/beta)/(gamma(1 + 1/beta))^2 - (1 + cv2))
}
k = optimize(fun, interval = c(0, 100), cv2 = cv2, maximum = FALSE)$minimum
lambda = xb/gamma(1 + 1/k)
k
lambda
```

Then, we are going to overlay the estimated pdf onto histogram.

```{r}
hist(ht, breaks = 20, freq = FALSE)
curve(dweibull(x, k, lambda), add = TRUE)
```

Overlay the estimated cdf onto the ecdf

```{r}
plot(ecdf(ht), cex = 0.1, col = "red", main = "", ylab = "Density")
curve(pweibull(x, k, lambda), add = TRUE)
legend('bottomright', legend = c("estimated cdf", "ecdf"), col = c("black", "red"), lty = 1, lwd = 2)
```

QQ plot

```{r}
random_sample <- ht
q_candidate <- qweibull

x <- q_candidate((1:200)/200, k, lambda)
y <- quantile(ht, probs = (1:200)/200)

tgsify::plotstyle(style = upright)
plot(x,y, asp = 1, xlab = "Estimated quantile", ylab = "Sample quantile")
abline(0,1)
```

Estimated median

```{r}
qweibull(0.5, k, lambda)
```

Median sampling distribution

```{r}
median <- c()
for (i in 1:1e5) {
  data <- rweibull(1e4, k, lambda)
  median <- c(median, median(data))
}

hist(median, breaks = 20, freq = FALSE)
```

Range of middle 95% of sampling distribution

```{r}
range = quantile(median, c(0.025, 0.975))
range[2] - range[1]
```

