---
title: "writeup"
output:
  html_document:
    df_print: paged
---

It is common in the analysis of biological data to log transform data representing concentrations or data representing dose response. In this project, I will operate the log transform to different distributions and let's see the results.

# Part 1

## Distribution 1

The first distribution we want to explore is the gamma distribution with shape parameter equal to 3 and scale parameter equal to 1.

Generate a figure of the PDF and CDF. Mark the mean and median in the figure.

```{r}
# Probability distribution function
x = seq(0, 20, 0.05)
y = dgamma(x, shape = 3, scale = 1)

plot(x, y, main="Gamma Probability Distribution Function", ylab = "Density", xlim = c(0, 20), ylim = c(0, 1), col = "red", type = "l", lwd = 2)
legend("topright", legend = paste("shape=", 3, "scale=", 1), lwd = 2, col = "red")
abline(v = 3*1, col = "blue")
text(x = 3, y = 0.8, "mean = 3", pos = 4)
abline(v = qgamma(0.5, shape = 3, scale = 1), col = "green")
text(x = 2.67, y = 0.6, "median \n= 2.67", pos = 2)
```

```{r}
# Cumulative distribution function
x = seq(0, 20, 0.05)
y = pgamma(x, shape = 3, scale = 1)

plot(x, y, main="Gamma Cumulative Distribution Function", ylab = "Cumulative Density", xlim = c(0, 20), ylim = c(0, 1), col = "red", type = "l", lwd = 2)
legend("right", legend = paste("shape=", 3, "scale=", 1), lwd = 2, col = "red")
abline(v = 3*1, col = "blue")
text(x = 3, y = 0.8, "mean = 3", pos = 4)
abline(v = qgamma(0.5, shape = 3, scale = 1), col = "green")
text(x = 2.67, y = 0.6, "median \n= 2.67", pos = 2)
```

Generate a figure of the PDF and CDF of the transformation Y = log(X) random variable. Mark the mean and median in the figure. You may use simulation or analytic methods in order find the PDF and CDF of the transformation.

I use the simulation method to find the PDF and CDF of the transformation.

```{r}
# Probability distribution function
set.seed(1)
x = rgamma(1e6, shape = 3, scale = 1)
y = log(x)
plot(density(y), main = "Gamma Probability Distribution Function(Log Transformation)",  xlab = "log(x)", ylim = c(0, 1), col = "red", type = "l", lwd = 2)
abline(v = mean(y), col = "blue")
text(x = mean(y), y = 0.8, "mean\n(blue line)", pos = 2)
abline(v = median(y))
text(x = median(y), y = 0.7, "median(black line)", pos = 4)
```

```{r}
# Cumulative distribution function
set.seed(2)
x = rgamma(1e6, shape = 3, scale = 1)
y = log(x)
plot(ecdf(y), main = "Gamma Cumulative Distribution Function(Log Transformation)", xlab = "log(x)", ylab = "Cumulative Density", ylim = c(0, 1), col = "red", lwd = 2)
abline(v = mean(y), col = "blue")
text(x = mean(y), y = 0.8, "mean\n(blue line)", pos = 2)
abline(v = median(y))
text(x = median(y), y = 0.7, "median(black line)", pos = 4)
```

Generate 1000 samples of size 100. For each sample, calculate the geometric and arithmetic mean. Generate a scatter plot of the geometic and arithmetic sample means. Add the line of identify as a reference line.

```{r}
x <- matrix(rgamma(1e5, shape = 3, scale = 1), nrow = 1000, ncol = 100)
geometric_mean <- function(x){
  exp(mean(log(x)))
}
geometric_mean1 <- apply(x, 1, geometric_mean)
arithmetric_mean1 <- apply(x, 1, mean)
plot(x = geometric_mean1, y = arithmetric_mean1)
abline(0, 1)
```

Generate a histogram of the difference between the arithmetic mean and the geometric mean

```{r}
difference <- arithmetric_mean1 - geometric_mean1
hist(difference, main = "Hisyogram of difference between arithmetic and geometric mean")
```

## Distribution 2

The second distribution we want to explore is the logarithmic normal distribution with $\mu = -1$ and $\sigma = 1$.

Generate a figure of the PDF and CDF. Mark the mean and median in the figure.

```{r}
# Probability distribution function
x = seq(0, 20, 0.05)
y = dlnorm(x, mean = -1, sd = 1)

plot(x, y, main="Logarithmic Normal Probability Distribution Function", ylab = "Density", xlim = c(0, 20), ylim = c(0, 2), col = "red", type = "l", lwd = 2)
legend("topright", legend = expression(paste(mu, " = -1", sigma, " = 1")), lwd = 2, col = "red")
abline(v = exp(1/2), col = "blue")
text(x = exp(1/2), y = 0.8, "mean = 1.65", pos = 4)
abline(v = exp(-1), col = "green")
text(x = exp(-1), y = 1.5, "median \n= 0.37", pos = 4)
```

```{r}
# Cumulative distribution function
x = seq(0, 20, 0.05)
y = plnorm(x, mean = -1, sd = 1)

plot(x, y, main="Logarithmic Normal Cumulative Distribution Function", ylab = "Cumulative Density", xlim = c(0, 20), ylim = c(0, 1), col = "red", type = "l", lwd = 2)
legend("right", legend = expression(paste(mu, " = -1", sigma, " = 1")), lwd = 2, col = "red")
abline(v = exp(1/2), col = "blue")
text(x = exp(1/2), y = 0.8, "mean = 1.65", pos = 4)
abline(v = exp(-1), col = "green")
text(x = exp(-1), y = 0.5, "median = 0.37", pos = 4)
```

Generate a figure of the PDF and CDF of the transformation Y = log(X) random variable. Mark the mean and median in the figure. You may use simulation or analytic methods in order find the PDF and CDF of the transformation.

I use the simulation method to find the PDF and CDF of the transformation.

```{r}
# Probability distribution function
set.seed(3)
x = rlnorm(1e6, mean = -1, sd = 1)
y = log(x)
plot(density(y), main = "Logarithmic Normal Probability Distribution Function(Log Transformation)",  xlab = "log(x)", ylim = c(0, 1), col = "red", type = "l", lwd = 2)
abline(v = mean(y))
text(x = mean(y), y = 0.8, "mean", pos = 2)
abline(v = median(y))
text(x = median(y), y = 0.7, "median", pos = 4)
```

```{r}
# Cumulative distribution function
set.seed(4)
x = rlnorm(1e6, mean = -1, sd = 1)
y = log(x)
plot(ecdf(y), main = "Logarithmic Normal Cumulative Distribution Function(Log Transformation)", xlab = "log(x)", ylab = "Cumulative Density", ylim = c(0, 1), col = "red", lwd = 2)
abline(v = mean(y))
text(x = mean(y), y = 0.8, "mean", pos = 2)
abline(v = median(y))
text(x = median(y), y = 0.7, "median", pos = 4)
```

Generate 1000 samples of size 100. For each sample, calculate the geometric and arithmetic mean. Generate a scatter plot of the geometic and arithmetic sample means. Add the line of identify as a reference line.

```{r}
x <- matrix(rlnorm(1e5, mean = -1, sd = 1), nrow = 1000, ncol = 100)
geometric_mean <- function(x){
  exp(mean(log(x)))
}
geometric_mean1 <- apply(x, 1, geometric_mean)
arithmetric_mean1 <- apply(x, 1, mean)
plot(x = geometric_mean1, y = arithmetric_mean1)
abline(0,1)
```

Generate a histogram of the difference between the arithmetic mean and the geometric mean

```{r}
difference <- arithmetric_mean1 - geometric_mean1
hist(difference, main = "Hisyogram of difference between arithmetic and geometric mean")
```

## Distribution 3

The third distribution we want to explore is the uniform distribution between 0 and 12.

Generate a figure of the PDF and CDF. Mark the mean and median in the figure.

```{r}
# Probability distribution function
x = seq(0, 12, 0.01)
y = dunif(x, 0, 12)

plot(x, y, main="Uniform Probability Distribution Function", ylab = "Density", xlim = c(0, 15), ylim = c(0, 1), col = "red", type = "l", lwd = 2)
legend("topright", legend = paste("min=", 0, "max=", 12), lwd = 2, col = "red")
abline(v = 6, col = "blue")
text(x = 6, y = 0.8, "mean = 6", pos = 4)
abline(v = 6, col = "green")
text(x = 6, y = 0.6, "median = 6", pos = 2)
```

```{r}
# Cumulative distribution function
x = seq(0, 12, 0.01)
y = punif(x, 0, 12)

plot(x, y, main="Uniform Cumulative Distribution Function", ylab = "Cumulative Density", xlim = c(0, 15), ylim = c(0, 1), col = "red", type = "l", lwd = 2)
legend("right", legend = paste("min=", 0, "max=", 12), lwd = 2, col = "red")
abline(v = 6, col = "blue")
text(x = 6, y = 0.8, "mean = 6", pos = 4)
abline(v = 6, col = "green")
text(x = 6, y = 0.6, "median = 6", pos = 2)
```

Generate a figure of the PDF and CDF of the transformation Y = log(X) random variable. Mark the mean and median in the figure. You may use simulation or analytic methods in order find the PDF and CDF of the transformation.

I use the simulation method to find the PDF and CDF of the transformation.

```{r}
# Probability distribution function
set.seed(5)
x = runif(1e6, 0, 12)
y = log(x)
plot(density(y), main = "Uniform Probability Distribution Function(Log Transformation)",  xlab = "log(x)", ylim = c(0, 1), col = "red", type = "l", lwd = 2)
abline(v = mean(y), col = "blue")
text(x = mean(y), y = 0.8, "mean(blue line)", pos = 2)
abline(v = median(y))
text(x = median(y), y = 0.7, "median(black line)", pos = 2)
```

```{r}
# Cumulative distribution function
set.seed(6)
x = runif(1e6, 0, 12)
y = log(x)
plot(ecdf(y), main = "Uniform Cumulative Distribution Function(Log Transformation)", xlab = "log(x)", ylab = "Cumulative Density", ylim = c(0, 1), col = "red", lwd = 2)
abline(v = mean(y), col = "blue")
text(x = mean(y), y = 0.8, "mean(blue line)", pos = 2)
abline(v = median(y))
text(x = median(y), y = 0.7, "median(black line)", pos = 2)
```

Generate 1000 samples of size 100. For each sample, calculate the geometric and arithmetic mean. Generate a scatter plot of the geometic and arithmetic sample means. Add the line of identify as a reference line.

```{r}
x <- matrix(runif(1e5, 0, 12), nrow = 1000, ncol = 100)
geometric_mean <- function(x){
  exp(mean(log(x)))
}
geometric_mean1 <- apply(x, 1, geometric_mean)
arithmetric_mean1 <- apply(x, 1, mean)
plot(x = geometric_mean1, y = arithmetric_mean1)
abline(0,1)
```

Generate a histogram of the difference between the arithmetic mean and the geometric mean

```{r}
difference <- arithmetric_mean1 - geometric_mean1
hist(difference, main = "Hisyogram of difference between arithmetic and geometric mean")
```

# Part 2

The arithmetric mean of c($X_1, X_2,..., X_n$) is ($X_1+X_2+...+X_n$)/n.

The geometric mean of c($X_1, X_2,..., X_n$) is $\sqrt[n]{X_1+X_2+...+X_n}$, and we can calculate the value by using exp(($log(X_1)+log(X_2)+...+log(X_n)$)/n).

I will use simulation method to show this rule.

```{r}
# Set the random number seed.
set.seed(7)
arithmetric_mean2 <- c()
geometric_mean2 <- c()
for (i in 1:1e4) {
# Generate uniform distribution random number and calculate their arithmetric mean and geometric mean separately.
  x <- runif(1e4, 0, 1e6)
  arithmetric_mean2 <- c(arithmetric_mean2, mean(x))
  geometric_mean2 <- c(geometric_mean2, exp(mean(log(x))))
}
# Plot the histogram of the difference between arithmetric mean and geometric mean.
hist(arithmetric_mean2 - geometric_mean2, main = "difference")
```

As the above graph shows, the histogram of difference has all the values larger than 0, so we can make a conclusion that the arithmetric mean is larger than the geometric mean.

# Part 3

I will use simulation and analytic method to show that log(E(x)) is always larger than E[log(x)].

```{r}
set.seed(8)
log_exp <- c()
exp_log <- c()
# Set a for loop to replicate the data generate process and verify if the conclusion is correct for all cases.
for (i in 1:1e4) {
# Generate uniform distribution random number.
  x <- runif(1e3, 0, 1e6)
# Calculate the values and store them into the variables.
  log_exp <- c(log_exp, log(mean(x)))
  exp_log <- c(exp_log, mean(log(x)))
}
# Plot a histogram to see the results.
hist(log_exp - exp_log, main = "difference")
```

As the above graph shows, the histogram of difference has all the values larger than 0, so we can make a conclusion that the log(mean(x)) is larger than the mean(log(x)).

Analytic method

As we all know, we can get the following result through the plot:

$e^{x} \geq (x+1)$

```{r}
x = seq(-3, 3, 0.02)
y1 = exp(x)
y2 = x + 1
plot(x, y1, type = "l")
lines(x, y2, type ="l")
```

Imagine a random variable y:

$$
E(e^y) = E(e^{y+E(y)-E(y)})\\

       = e^{E(y)}E(e^{y-E(y)})\\
       
       \geq e^{E(y)}E(y-E(y)+1)\\
       
E(e^y)\geq e^{E(y)}\\

if\quad y = lnx\\

E(x)\geq e^{E(lnx)}\\

Do logarithmic processing for the both sides.\\

ln(E(x))\geq E(ln(x))

$$
Then, we can get the result the log(mean(x)) no less than mean(log(x)).