---
title: "writeup"
author: "Weixi Chen"
date: "2021/9/24"
output: html_document
---

# Background

The world series is a best-of-7 match-up between the champions of the American and National Leagues of Major League Baseball. Now, we are going to focus on an imagine situation. Here are two assumptions:

1. Suppose that the Braves and the Yankees are teams competing in the World Series.
2. Suppose that in any given game, the probability that the Braves win is $P_{B}$ and the probability that the Yankees win is $ P_{Y}=1-P_{B}$.

# Question 1

What is the probability that the Braves win the World Series given that $P_{B}$ = 0.55?

answer:

```{r}
# We can use the negative binomial distribution to tackle this problem. "pnbinom" is a 
# function in which we could find out the cumulative distribution values at any point. The # cumulative distribution value at "a" means the sum of probability at points equal to or 
# less than "a". In this question, We are going to explore the probability that the Braves # win the World Series, which means that the Braves needs to win 4 games during the 7 
# total games. Thus, we need to calculate sum of the probability that Braves win 4 games 
# and lose 1 game, win 4 games and lose 2 games, and win 4 games and lose 3 games.
p = pnbinom(3, 4, 0.55)
p
```

The probability that the Braves win the World Series given that $P_{B}$ = 0.55 is 0.6082878.

# Question 2

What is the probability that the Braves win the World Series given that $P_{B}=x$? This will be a figure (see below) with PB on the x-axis and P(Braves win World Series) on the y-axis

This is a question that is quite similar as the question1, we just need to change the fixed probability 0.55 into x, assign x with different values, store the results, and plot the graph.
```{r}
# Set p_b from 0.02 to 0.98 with a interval of 0.02.
p_b <- seq(0.02, 0.98, 0.02)
# Define p_bw to store the probability that the Braves win the World Series.
p_bw <- c()
# Set a for loop to calculate the p_bw when p_b changes and save the results.
for (x in p_b) {
  p_bw = c(p_bw, pnbinom(3, 4, x))
}
# Plot the graph and show it.
p <- plot(x = p_b, 
          y = p_bw, 
          xlim = c(0.5, 1),
          main = "Probability of the Braves winning the World Series", 
          xlab = "Probability of the Braves winning a head-to-head matchup", 
          ylab = "Pr(Win World Series)")
p
```

From the graph, we can find out the probability that the Braves win the World Series goes up when $P_{B}$ goes up; what's more, the speed of P(Braves win World Series)'s increase declines gradually.

# Question 3

Suppose one could change the World Series to be best-of-9 or some other best-of-X series. What is the shortest series length so that P(Braves win World Series|$P_{B}$=0.55)$\ge$0.8

```{r}
# We set the for loop from x=1 to x=100
for (x in 1:100) {
# Test if P(Braves win World Series) >= 0.8 under the condition that p_b = 0.55
  if(pnbinom((x-1)/2, (x+1)/2, 0.55) >= 0.8){
# Print the stop point x, which is the shortest series length that meet the condition and the corresponding probability P(Braves win World Series|p_b = 0.55)
    print(x)
    print(pnbinom((x-1)/2, (x+1)/2, 0.55))
# End the for loop if the x is found.
    break
  }
}
```

If one could change the World Series to be best-of-9 or some other best-of-X series. The shortest series length so that P(Braves win World Series|$P_{B}$=0.55)$\ge$0.8 is 71, and the P(Braves win World Series|$P_{B}$=0.55) = 0.8017017.

# Question 4

What is the shortest series length so that P(Braves win World Series|$P_{B}$ = x)$\ge$0.8? This will be a figure (see below) with PB on the x-axis and series length is the y-axis.

```{r}
#Set the probability of p_b and the y_shortest to store the shortest series length when p_b changes.
p_b <- seq(0.02, 0.98, 0.02)
y_shortest <- c()
# We set the for loop from x in p_b and y from 1 to 2000.
for (x in p_b) {
  for (y in 1:2000) {
#Check if the stopping rule is met.
    if(pnbinom((y-1)/2, (y+1)/2, x) >= 0.8){
      break
    }
  }
#Store the shortest series length.
  y_shortest <- c(y_shortest, y)
}
# Plot the graph and show it.
p1 <- plot(x = p_b, 
           y = y_shortest, 
           xlim = c(0.5, 1),
           main = "Shortest Series so that P(Win WS given p) >= 0.8", 
           xlab = "Probability of the Braves winning a head-to-head matchup", 
           ylab = "Series length")
p1
```

From the graph we can find out that the shortest series length declines as the probability of the Braves winning a head-to-head match goes up, and the speed for its decline goes down gradually. In the end, the shortest series length stablises at 1 game when the probability of the Braves winning a head-to-head matchup is lager than 0.8.

# Question 5

Calculate P($P_{B}$ =0.55|Braves win World Series in 7 games) under the assumption that either $P_{B}$ = 0.55 or $P_{B}$ = 0.45. Explain your solution.

First, we know that $P(P_{B} = 0.55)$ = 1/2.
And, based on the Bayes Rule, we can also know that:

P($P_{B}$ =0.55|Braves win World Series in 7 games) = P(Braves win World Series in 7 games|$P_{B}$ =0.55)*P($P_{B}$ = 0.55)/P(Braves win World Series in 7 games)

We can know that P(Braves win World Series in 7 games|$P_{B}$ =0.55) = 0.1667701.

```{r}
p3 = dnbinom(3, 4, 0.55)
p3
```


Then, we need to calculate P(Braves win World Series in 7 games), which is equal to:
(P(Braves win World Series in 7 games|$P_{B}$ =0.55) + P(Braves win World Series in 7 games|$P_{B}$ =0.45))/2. So, we just need to calculate P(Braves win World Series in 7 games|$P_{B}$ =0.45).
```{r}
p2 = dnbinom(3, 4, 0.45)
p2
```

The P(Braves win World Series in 7 games|$P_{B}$ =0.45) is p2 = 0.1364483

```{r}
p_wanted = (0.1667701*0.5)/(0.5*(0.1667701+0.1364483))
p_wanted
```

As a result, P($P_{B}$ =0.55|Braves win World Series in 7 games) under the assumption that either $P_{B}$ = 0.55 or $P_{B}$ = 0.45 is equal to 0.5499999.